---
title: "Flights"
author: "Eddie Liu"
date: "November 6, 2019"
output: html_document
---

#Which Locations and Airlines experience more delays?

##Delays by Airlines
```{r}
library(dplyr)
library(ggplot2)
mydata <- read.csv('flights.csv')
dim(mydata)


#Absolute value of Departure and Arrival delays
mydata$DEPARTURE_DELAY <- abs(mydata$DEPARTURE_DELAY)
mydata$ARRIVAL_DELAY <- abs(mydata$ARRIVAL_DELAY)


#Summary table of amount of Airlines
table(mydata$AIRLINE)



#Average Departure Delay Times by Airline
AIRLINE_dep_delay <- mydata %>% group_by(AIRLINE) %>%
           summarise(
             AverageDelay = mean(DEPARTURE_DELAY, na.rm = TRUE)) %>% arrange(desc(AverageDelay))

AIRLINE_dep_delay 

par(mfrow=c(1,2)) 
ggplot(data = AIRLINE_dep_delay, aes(x=reorder(AIRLINE, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='steelblue') + xlab('Airline') + ylab('Average Delay') + ggtitle('Average Departure Delay Times by Airline') + theme(plot.title = element_text(hjust = 0.5))

```


```{r}
#Average Arrival Delay Times by Airline
AIRLINE_arr_delay <- mydata %>% group_by(AIRLINE) %>%
            summarise(
             AverageDelay = mean(ARRIVAL_DELAY, na.rm = TRUE))

AIRLINE_arr_delay

ggplot(data = AIRLINE_arr_delay, aes(x=reorder(AIRLINE, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='orange') + xlab('Airline') + ylab('Average Delay') + ggtitle('Average Arrival Delay Times by Airline') + theme(plot.title = element_text(hjust = 0.5))
```



##Delays by Location
###Origin Airport
```{r}
#Average Departure Delay Times by Top 5 Origin Airport
AIRLINE_dep_delay <- mydata %>% group_by(ORIGIN_AIRPORT) %>%
           summarise(
             n = n(),
             AverageDelay = mean(DEPARTURE_DELAY, na.rm = TRUE)) %>% arrange(desc(n))

top_5_airlines <- head(AIRLINE_dep_delay,5)
top_5_airlines

par(mfrow=c(1,2)) 
ggplot(data = top_5_airlines, aes(x=reorder(ORIGIN_AIRPORT, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='steelblue') + xlab('Airline') + ylab('Average Delay') + ggtitle('Top 5 Average Departure Delay Times by Origin Airport') + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#Average Arrival Delay Times by Top 5 Origin Airport
AIRLINE_arr_delay <- mydata %>% group_by(ORIGIN_AIRPORT) %>%
           summarise(
             n = n(),
             AverageDelay = mean(ARRIVAL_DELAY, na.rm = TRUE)) %>% arrange(desc(n))

top_5_airlines <- head(AIRLINE_arr_delay,5)
top_5_airlines


ggplot(data = top_5_airlines, aes(x= reorder(ORIGIN_AIRPORT, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='steelblue') + xlab('Airline') + ylab('Average Delay') + ggtitle('Top 5 Average Arrival Delay Times by Origin Airport') + theme(plot.title = element_text(hjust = 0.5))
```


###Destination Airport
```{r}
#Average Departure Delay Times by Top 5 Destination Airport
AIRLINE_dep_delay <- mydata %>% group_by(DESTINATION_AIRPORT) %>%
           summarise(
             n = n(),
             AverageDelay = mean(DEPARTURE_DELAY, na.rm = TRUE)) %>% arrange(desc(n))

top_5_airlines <- head(AIRLINE_dep_delay,5)
top_5_airlines

par(mfrow=c(1,2)) 
ggplot(data = top_5_airlines, aes(x=reorder(DESTINATION_AIRPORT, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='steelblue') + xlab('Airline') + ylab('Average Delay') + ggtitle('Top 5 Average Departure Delay Times by Destination Airport') + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#Average Arrival Delay Times by Top 5 Destination Airport
AIRLINE_arr_delay <- mydata %>% group_by(DESTINATION_AIRPORT) %>%
           summarise(
             n = n(),
             AverageDelay = mean(ARRIVAL_DELAY, na.rm = TRUE)) %>% arrange(desc(n))

top_5_airlines <- head(AIRLINE_arr_delay,5)
top_5_airlines


ggplot(data = top_5_airlines, aes(x= reorder(DESTINATION_AIRPORT, AverageDelay), y = AverageDelay)) + geom_bar(stat = 'identity', fill='steelblue') + xlab('Airline') + ylab('Average Delay') + ggtitle('Top 5 Average Arrival Delay Times by Destination Airport') + theme(plot.title = element_text(hjust = 0.5))
```



#Additional visualizations + Research airlines and see if their reputation matches what the data says (ex: spirit is not that good so makes sense for more delays)
```{r}

```




#Relationship between Departure Delay and Arrival Delay
```{r}
#Relationship between Departure Delay and Arrival Delay
cor.test(mydata$DEPARTURE_DELAY, mydata$ARRIVAL_DELAY)

plot(mydata$DEPARTURE_DELAY, mydata$ARRIVAL_DELAY, xlab = 'Departure Delay', ylab = 'Arrival Delay', main = "Association between Departure Delay and Arrival Delay")
```

###There appears to be a definite positive correlation between the Departure Delay time and Arrival Delay time.

##Find out average prices for the 2 airlines with the top airline delays and see if there's a correlation? Correlation between price and delays?





